Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems

### 2024-01-10 2024pipit/section/30-Core.tex
core/causality lemmas

% Recursive expressions are the essence of computation, so their properties are particularly important.
Our semantics for a recursive expression $\xrec{x}{e}$ performs one step of recursion by substituting $x$ for the recursive expression.
An alternative semantics would be to have the environment outside the semantics invent a stream $V$ such that if we extend the streaming history with $x \mapsto V$, then $e$ evaluates to $V$ itself.
\TODO{XYZ}
Finally, we also have that evaluation of a causal recursive expression is not affected by the most recent recursive value:
\begin{theorem}[bigstep-rec-causal]
  For a streaming history $\Sigma; \sigma$ and a causal recursive expression $\xrec{x}{e}$, the evaluation of $e$ does not depend on the most recent value of $x$ in the streaming history: if $(\bigstep{\Sigma; \sigma}{e}{v})$, then substituting $\sigma[x]$ with any value $v'$ results in the same value: $(\bigstep{\Sigma; \sigma[x := v']}{e}{v})$.
\end{theorem}

>>>>

The semantics in \autoref{f:core-bigstep} for a recursive expression $\xrec{x}{e}$ performs one step of recursion by substituting $x$ for the recursive expression.
An alternative semantics would be to have the environment outside the semantics invent a stream $V$ such that if we extend the streaming history with $x \mapsto V$, then $e$ evaluates to $V$ itself.
The above substitution theorem can be used to show that these two semantics are equivalent.
Thanks to causality, we can additionally show that, when evaluating $e$ with $x \mapsto V$, the most recent value in $V$ does not affect the result.
This fact can be used to ``seed'' evaluation by starting with an arbitrary value:
\begin{theorem}[bigstep-rec-causal]
  For a streaming history $\Sigma; \sigma$ and a causal recursive expression $\xrec{x}{e}$, if $(\bigstep{\Sigma; \sigma}{e}{v})$, then updating $\sigma[x]$ with any value $v'$ results in the same value: $(\bigstep{\Sigma; \sigma[x := v']}{e}{v})$.
\end{theorem}
